---
title: "Pitch: Geographical information system for health-related outcomes"
author: "Dr. Mathieu Licht"
date: "2024-06-16"
output: slidy_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(leaflet)
library(tidyverse)
```

## Tool overview
- **Aim:** detect spatial patterns of health-related outcomes
- **Data source:** Centers for Disease Control and Prevention (CDC)
- **Demo:** restricted to one region, here: California (USA)

## Health related outcomes for analysis
The tool places markers on the map indicating the crude (i.e., unadjusted) prevalence of various outcomes, like: 

- Diseases (e.g., chronic kidney disease)
- Risk factors (e.g., smoking)
- Preventive measures (e.g., cancer screening)
- Disabilities 
- Health status

Prevalences are stratified across census tracts.

## Concept (map)
Example for 2 locations: diabetes prevalence. 

- More blueish colors and bigger circles indicate a higher prevalence
- Click on a marker to learn more about the location (population size, prevalence rate in %, FIPS code).  

```{r show-map, echo = FALSE}
my_map <- leaflet() %>%
  addTiles()  %>% 
  addCircleMarkers(
    lat = 37.86754, 
    lng = -122.2318,
    radius = 1.5 * 8.4, 
    weight = 1, 
    fillColor = "#DCE319", 
    fillOpacity = 0.7,
    popup = "Alameda, Tract FIPS 6001400100 <br> Total Population: 2937 <br> Mean crude prevalence: 8.4%"
  ) %>% 
  addCircleMarkers(
    lat = 37.81204, 
    lng = -122.2824,
    radius = 1.5 * 13, 
    weight = 1, 
    fillColor = "#2D708E", 
    fillOpacity = 0.7,
    popup = "Alameda, Tract FIPS 6001402400 <br> Total Population: 2351 <br> Mean crude prevalence: 13%"
  ) 

my_map
```

## Additional descriptive statistics 
For each prevalence, the tool will display summary statistics, as shown below:


```{r randomexample, echo = FALSE}
set.seed = 123
random_prevalence <- rnorm(2000, 8, sd = 2)

the_mean <- round(mean(random_prevalence), digits = 1)
the_sd   <- round(sd(random_prevalence), digits = 1)
the_ci_lower <- round(the_mean  - 1.96 * the_sd / sqrt (length(random_prevalence)), digits = 1)
the_ci_upper <- round(the_mean  + 1.96 * the_sd / sqrt (length(random_prevalence)), digits = 1)

the_median <- round(median(random_prevalence, na.rm = TRUE), digits = 1)
the_q1 <- round(quantile(random_prevalence, prob=0.25)[[1]], digits = 1)
the_q3 <- round(quantile(random_prevalence, prob=0.75)[[1]], digits = 1)

```

**Prevalence (%):**

- **Mean (SD):** `r the_mean` (`r the_sd`)
- **95% Confidence interval:** `r the_ci_lower` to `r the_ci_upper`
- **Median (quartile 1; 3):** `r the_median` (`r the_q1`; `r the_q3`)  

```{r histogram, echo = FALSE, warning = FALSE, fig.height = 4}
ggplot(data = data.frame(x = random_prevalence), aes(x = x))  + 
            geom_histogram(color = "black" , fill = "#19d68f", show.legend = FALSE, binwidth = 2) +
            xlab("Crude prevalence (%)") +
            ylab("Count")+ 
            theme_bw() 
```

## Other features

- Switch between 36&nbsp;health-related outcomes on the fly
- Customize circle marker options (size, color, scale to population size or prevalence)
- Filter for ranges (population, prevalences) 
- Customize bin widths of histograms

